---
title: "Assignment 2"
---


```{r setup, include=FALSE}
library(ggplot2)
library(gtrendsR)
```


# Analysis by Downloading Google Trends Data

```{r}
# analyze the data after cleaning it
election <- read.csv('/Users/bepi36/Library/CloudStorage/OneDrive-TheUniversityofTexasatDallas/UTD/Fall2025/Data collection/assignment_2/election.csv', skip = 1)
names(election) <- c("date", "hits")
election$hits <- as.character(election$hits)
election$hits <- ifelse(election$hits == "<1", 0, as.numeric(election$hits))


ggplot(election, aes(x = date, y = hits, group = 1)) +
  geom_line()

# kamala data

kamala <- read.csv('/Users/bepi36/Library/CloudStorage/OneDrive-TheUniversityofTexasatDallas/UTD/Fall2025/Data collection/assignment_2/kamala.csv', skip = 1)
names(kamala) <- c("date", "hits")
kamala$hits <- as.character(kamala$hits)
kamala$hits <- ifelse(kamala$hits == "<1", 0, as.numeric(kamala$hits))

ggplot(kamala, aes(x = date, y = hits, group = 1)) +
  geom_line()

# trump data 

trump <- read.csv('/Users/bepi36/Library/CloudStorage/OneDrive-TheUniversityofTexasatDallas/UTD/Fall2025/Data collection/assignment_2/trump.csv', skip = 1)
names(trump) <- c("date", "hits")
trump$hits <- as.character(trump$hits)
trump$hits <- ifelse(trump$hits == "<1", 0, as.numeric(trump$hits))

ggplot(trump, aes(x = date, y = hits, group = 1)) +
  geom_line()

```



The above code relies on the 3 different CSV files downloaded from Google Trends. The data is not clean and not properly formatted to be analyzed in R. As such, further data cleaning was necessary to visualize the data. Based on the graphs, we can see the trends for each term. For election, it is clear that American election seem to dominate searches with a seasonality of 4 years. Trump and Kamala present similar matches in terms of search trends, with spikes preceding elections in 2016, 2020 and 2024. Trump has higher search interest, especially for the 2016 and 2020 elections, showcasing his importance as a political candidate.

# Analysis by gtrendsR

```{r}
# analyze with gtrendsR

# class code -----

## EPPS 6302 Methods of Data Collection and Production
## Google Trends with R

TrumpHarrisElection = gtrends(c("Trump","Harris","election"), onlyInterest = TRUE, geo = "US", gprop = "web", time = "today+5-y", category = 0, ) # last five years
the_df=TrumpHarrisElection$interest_over_time
plot(TrumpHarrisElection)
tg = gtrends("tariff", time = "all")

```

Utilizing the gtrendsR package yields similar results as the manual download from google trends. Differences stand in how streamlined the process is, and the data being formatted in an R-friendly way. Substantive results are identical.
